---
title: "AcademyAwards"
author: "Emely P."
date: "2026-05-16"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
#import library
```{r}
library(ggplot2)
library(dplyr)
library(stringr)
library(readr)
library(tidyr)
library(reshape2)
library(jsonlite)
library(igraph)
library(tidygraph)
library(ggraph)
library(visNetwork)
library(scales)
```


# import csv
```{r}
data <- read_csv("~/Downloads/film_details_v3.csv")
```
#Frequency Count of every unique value in column "director" 
```{r}
table(data$director)
```
#Sort by most frequent
```{r}
sort(table(data$director), decreasing = TRUE)
```
#Plot in bar graph
```{r}
barplot(table(data$director))
```
#plot directors with more than 1 nomination
```{r}
dir_counts <- sort(table(data$director), decreasing = TRUE)
dir_counts_filtered <- dir_counts[dir_counts > 1]

barplot(dir_counts_filtered,
        las = 2,
        col = "navy",
        main = "Frequent Directors",
        ylab = "Number of Movies")
```
#Change orientation, names on y-axis
```{r}
barplot(dir_counts_filtered,
        horiz = TRUE,
        col = "navy",
        main = "Director Nominations Count",
        xlab = "Number of Movies",
        las = 1,
        space = 0.5)
```
#change bar graph so directors names are visable
```{r}
max_chars <- max(nchar(names(dir_counts_filtered)))
par(mar = c(5, max_chars * 0.7, 4, 2))

barplot(dir_counts_filtered,
        horiz = TRUE,
        col = "navy",
        main = "Director's Film Nomination Count",
        xlab = "Number of Movies",
        las = 1,
        space = 0.5)
```
# Total number of directors with more than 1 nomination
```{r}
dir_counts <- table(data$director)
sum(dir_counts > 1)
```
#Table of director counts
```{r}
dir_counts <- table(data$director)
table(dir_counts)
```
#Plot of director counts
```{r}
barplot(table(dir_counts),
        col = "navy",
        main = "Director Frequency Distribution",
        xlab = "Number of Nominations",
        ylab = "Number of Directors")
```
#Plot of director counts with y-axis going to 200
```{r}
freq <- table(dir_counts)

bp <- barplot(freq,
              col = "navy",
              main = "Director Frequency Distribution",
              xlab = "Number of Nominations",
              ylab = "Number of Directors",
              ylim = c(0, 200))   # force y-axis to 200
```
#add counts on top of bar
```{r}
bp <- barplot(freq,
              col = "navy",
              main = "Director Frequency Distribution",
              xlab = "Number of Nominations",
              ylab = "Number of Directors",
              ylim = c(0, 200))   # force y-axis to 200
text(x = bp,
     y = freq,
     labels = freq,
     pos = 3,        
     cex = 0.8)
```
#Cast frequency
```{r}
all_cast <- unlist(data[paste0("Cast ", 1:15)])
all_cast <- na.omit(all_cast)
cast_counts <- table(all_cast)
cast_counts
```
#Sort cast by frequency
```{r}
sort(cast_counts, decreasing = TRUE)
```
#Cast frequency table
```{r}
cast_counts <- table(all_cast)
cast_counts_sorted <- sort(cast_counts, decreasing = TRUE)
cast_counts_sorted[cast_counts_sorted > 1]
```
#Plot cast distribution
```{r}
cast_freq_dist <- table(cast_counts)
cast_freq_dist
barplot(cast_freq_dist,
        col = "darkgreen",
        main = "Distribution of Actor Frequencies",
        xlab = "Number of Appearances",
        ylab = "Number of Actors")
```
#Add cast counts on of bars
```{r}
cast_freq_dist <- table(cast_counts)
bp2 <- barplot(cast_freq_dist,
              col = "darkgreen",
              main = "Distribution of Actor Frequencies",
              xlab = "Number of Appearances",
              ylab = "Number of Actors",
              ylim = c(0, 4000))
text(x = bp2,
     y = cast_freq_dist,
     labels = cast_freq_dist,
     pos = 3,      # above bars
     cex = 0.8)
```
#Nominated Actor/Actress Counts from Nominee Column
```{r}
AA_final <- read_csv("~/data/628_AcademyAwards/AA_final.csv")
nominee_counts <- table(AA_final$Nominee)
nominee_counts
```
#Nominee counts for nominations only of Actors/Actress in Nominee column
```{r}
data_role <- AA_final[grepl("role", AA_final$Category, ignore.case = TRUE), ]
nominee_counts <- table(data_role$Nominee)
nominee_counts
```
#Nominee counts for nomination only of Actors/Actress in Nominee Column
```{r}
data_role <- AA_final[grepl("role", AA_final$Category, ignore.case = TRUE), ]
nominee_counts <- table(data_role$Nominee)
nominee_counts
```


#Remove actress in leading role entry, and sort
```{r}
nominee_counts <- table(AA_final$Nominee)
nominee_counts <- nominee_counts[!grepl("ACTRESS IN A LEADING ROLE",
                                        names(nominee_counts),
                                        ignore.case = TRUE)]
sort(nominee_counts, decreasing = TRUE)
```
#Count nominee frequency bar chart
```{r}
barplot(nominee_counts,
        horiz = TRUE,
        col = "pink",
        main = "Nominee Frequency",
        xlab = "Number of Nominations",
        las = 1)
```
#Nominee frequency distribution
```{r}
freq_dist <- table(nominee_counts)
freq_dist
barplot(freq_dist,
        col = "pink",
        xlab = "Number of Actor/Actress Nominations",
        ylab = "Number of Nominees",
        main = "Nominee Frequency Distribution")
```
#Add labels to bar chart
```{r}
bp3 <- barplot(freq_dist,
              col = "pink",
              ylim = c(0, max(freq_dist) * 1.1),
              xlab = "Number of Nominations",
              ylab = "Number of Nominees",
              main = "Nominee Frequency Distribution")

text(bp3,
     freq_dist,
     labels = freq_dist,
     pos = 3,
     cex = 0.8)
```
#Winner Frequency
```{r}
winners_role <- AA_final[
  AA_final$Award_Status == "Winner" &
  grepl("ROLE", AA_final$Category, ignore.case = TRUE),
]
winner_role_counts <- table(winners_role$Nominee)
sort(winner_role_counts, decreasing = TRUE)
```
#Plot Winner Frequency
```{r}
win_distribution <- table(winner_role_counts)
win_distribution
barplot(win_distribution,
        col = "chocolate",
        main = "Win Distribution",
        xlab = "Number of Wins",
        ylab = "Number of Winners")
```
#Add labels to winner frequency
```{r}
bp4 <- barplot(win_distribution,
              col = "chocolate",
              ylim = c(0, max(win_distribution) * 1.1),
              main = "Win Distribution",
              xlab = "Number of Wins",
              ylab = "Number of Winners")

text(bp4,
     win_distribution,
     labels = win_distribution,
     pos = 3,
     cex = 0.8)
```
#Directors with 4 or more associated nominations
```{r}
director_counts <- table(data$director)
topdirectors <- director_counts[director_counts >= 4]
topdirectors
```
#table of top directors
```{r}
data.frame(
  Director = names(topdirectors),
  Films = as.numeric(topdirectors)
)
```
#Filter top direcs in data.csv
```{r}
top_directors <- unique(data$director)
data_top <- data[data$director %in% top_directors, ]
```
#Verify count
```{r}
nrow(data_top)
unique(data_top$director)
```
#Cast List
```{r}
all_cast <- unlist(data[paste0("Cast ", 1:15)])
all_cast <- trimws(na.omit(all_cast))
all_cast <- all_cast[all_cast != ""]
unique_cast <- unique(all_cast)
```
#Split AA_Final Status
```{r}
all_people <- unique(AA_final$Award_Status)
winners <- unique(AA_final$Nominee[AA_final$Award_Status == "Winner"])
nominees <- unique(AA_final$Nominee[AA_final$Award_Status == "Nominee"])
```
#Cast not nominated
```{r}
all_award_people <- unique(AA_final$Award_Status)
not_nominated <- setdiff(unique_cast, all_award_people)
```
#Final Counts
```{r}
data.frame(
  Category = c("Won", "Nominated", "Not nominated"),
  Count = c(
    length(winners),
    length(nominees),
    length(not_nominated)
  )
)
```
# Change data.csv title column to film
```{r}
names(data)[names(data) == "title"] <- "Film"
```
#film-details import, change title to Film, and reformat
```{r}
film_details <- read_csv("~/Downloads/film_details.csv")
names(film_details)[names(film_details) == "title"] <- "Film"
cast <- film_details %>%
  separate_rows(top_cast, sep = ",") %>%
  mutate(Actor = trimws(top_cast)) %>%
  select(Film, director, Actor)
```
#Join cast & data role 
```{r}
combined <- cast %>%
  left_join(data_role, by = c("Film", "Actor" = "Nominee"))
```
#Stacked categories, Winner, Nominee, Non-Nominee
```{r}
combined <- combined %>%
  mutate(Status = case_when(
    Award_Status == "Winner" ~ "Winner",
    Award_Status == "Nominee" ~ "Nominee",
    TRUE ~ "Non-Nominee"
  ))
```
#plot stacked bar chart
```{r}
ggplot(combined, aes(x = director, fill = Status)) +
  geom_bar() +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
ggsave("plot2.png", width = 14, height = 8, dpi = 300)
```
#Switch x & y axis
```{r}
ggplot(combined, aes(x = director, fill = Status)) +
  geom_bar() +
  
  coord_flip() +
  
  theme_minimal() +
  
  theme(
    axis.text.y = element_text(size = 6),
    plot.margin = margin(10, 20, 20, 20)
  )

ggsave("plot2.png", width = 14, height = 14, dpi = 300)
```
#Filter director list
```{r}
film_filtered <- cast %>%
  filter(director %in% data$director)
```
#combined award data
```{r}
combined <- film_filtered %>%
  left_join(data_role, by = c("Film", "Actor" = "Nominee"))
```

#category list
```{r}
combined <- combined %>%
  mutate(Status = case_when(
    Award_Status == "Winner" ~ "Winner",
    Award_Status == "Nominee" ~ "Nominee",
    TRUE ~ "Non-Nominee"
  ))
```
#Plot stacked bar chart
```{r}
ggplot(combined, aes(x = director, fill = Status)) +
  geom_bar() +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

#top 20 directors
```{r}
top_directors <- combined %>%
  count(director, sort = TRUE) %>%
  slice_head(n = 20) %>%
  pull(director)

combined_top <- combined %>%
  filter(director %in% top_directors)

ggplot(combined_top, aes(x = reorder(director, -table(director)[director]), 
                         fill = Status)) +
  geom_bar() +
  theme_minimal() +
  labs(x = "Director", y = "Count") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
#Change order status
```{r}
combined$Status <- factor(
  combined$Status,
  levels = c("Winner","Nominee","Non-Nominee")
)
```
#Plot stacked bar chart
```{r}
top_directors <- combined %>%
  count(director, sort = TRUE) %>%
  slice_head(n = 20) %>%
  pull(director)

combined_top <- combined %>%
  filter(director %in% top_directors)

ggplot(combined_top, aes(x = reorder(director, -table(director)[director]), 
                         fill = Status)) +
  geom_bar() +
  theme_minimal() +
  labs(x = "Director", y = "Count") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

#Percentage stacked bar chart
```{r}
ggplot(combined_top, aes(x = director, fill = Status)) +
  geom_bar(position = "fill") +
  scale_y_continuous(labels = scales::percent) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
#count stacked bar chart 
```{r}
ggplot(combined_top, aes(x = director, fill = Status)) +
  geom_bar() +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(y = "Count", x = "Director")
```
summary table
```{r}
label_data <- combined_top %>%
  group_by(director, Status) %>%
  summarise(n = n(), .groups = "drop") %>%
  group_by(director) %>%
  mutate(total = sum(n))
```
#labels
```{r}
label_data <- label_data %>%
  mutate(label = n)
```
#stacked bar chart with labels
```{r}
ggplot(label_data, aes(x = director, y = n, fill = Status)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = label),
            position = position_stack(vjust = 0.5),
            size = 2) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(y = "Count", x = "Director")
```
#Director film count
```{r}
director_film_count_all <- combined %>%
  group_by(director) %>%
  summarise(film_count = n_distinct(Film))
print(director_film_count)
```
#top 20 director film count
```{r}
director_film_count <- combined_top %>%
  group_by(director) %>%
  summarise(film_count = n_distinct(Film))
print(director_film_count)
```
#All Actor/Actress frequency
```{r}
actor_frequency_all <- combined %>%
  group_by(Actor) %>%
  summarise(Frequency = n(), .groups = "drop") %>%
  arrange(desc(Frequency))
print(actor_frequency_all, n = Inf)
```

# Actor/Actress frequency in top 20 directors
```{r}
actor_frequency <- combined_top %>%
  group_by(Actor) %>%
  summarise(Frequency = n(), .groups = "drop") %>%
  arrange(desc(Frequency))
print(actor_frequency, n = Inf)
```
# Heat map code
#All Actor-Director Counts
```{r}
actor_director_all <- combined %>%
  group_by(director, Actor) %>%
  summarise(Frequency = n(), .groups = "drop")
```

#Top 20 Actor-Directo Counts
```{r}
actor_director <- combined_top %>%
  group_by(director, Actor) %>%
  summarise(Frequency = n(), .groups = "drop")
```

#Plot Heat Map All 
```{r}
ggplot(actor_director_all, aes(x = director, y = Actor, fill = Frequency)) +
  geom_tile() +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 6),
    axis.text.y = element_text(size = 6)
  ) +
  scale_fill_gradient(low = "white", high = "steelblue") +
  labs(title = "Actor–Director Collaboration Heatmap",
       x = "Director",
       y = "Actor",
       fill = "Frequency")
```
#Heat map of top 20 
```{r}
ggplot(actor_director, aes(x = director, y = Actor, fill = Frequency)) +
  geom_tile() +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 6),
    axis.text.y = element_text(size = 6)
  ) +
  scale_fill_gradient(low = "white", high = "steelblue") +
  labs(title = "Actor–Director Collaboration Heatmap",
       x = "Director",
       y = "Actor",
       fill = "Frequency")
```
# Heat Map with Actors with 2 or more counts (in all actors)
```{r}
valid_actors_all <- combined %>%
  count(Actor) %>%
  filter(n >= 2) %>%
  pull(Actor)

actor_director_filtered_all <- combined %>%
  filter(Actor %in% valid_actors_all) %>%
  count(director, Actor, name = "Frequency")

full_grid_all <- expand.grid(
  director = unique(actor_director_filtered_all$director),
  Actor = unique(actor_director_filtered_all$Actor)
)

actor_director_filtered_all <- full_grid_all %>%
  left_join(actor_director_filtered_all,
            by = c("director", "Actor")) %>%
  mutate(Frequency = ifelse(is.na(Frequency), 0, Frequency))

ggplot(actor_director_filtered_all, aes(x = director, y = Actor, fill = Frequency)) +
  geom_tile() +
  
  theme_minimal() +
  
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 6),
    axis.text.y = element_text(size = 6)
  ) +
  
  scale_fill_gradientn(
  colors = c("white","grey","blue4", "green3", "gold", "red"),
  values = scales::rescale(c(0, 1, 2, 3, 4, 5))
) +
  geom_tile(color = "grey90")+
  
  labs(
    title = "All Actor–Director Heatmap (Actors with ≥2 Appearances)",
    x = "Directors",
    y = "Actors",
    fill = "Frequency"
  )
```
# Heat map with actors with 4 or more counts in top 20 directors
```{r}
valid_actors <- combined_top %>%
  count(Actor) %>%
  filter(n >= 4) %>%
  pull(Actor)

actor_director_filtered <- combined_top %>%
  filter(Actor %in% valid_actors) %>%
  count(director, Actor, name = "Frequency")

full_grid <- expand.grid(
  director = unique(actor_director_filtered$director),
  Actor = unique(actor_director_filtered$Actor)
)

actor_director_filtered <- full_grid %>%
  left_join(actor_director_filtered,
            by = c("director", "Actor")) %>%
  mutate(Frequency = ifelse(is.na(Frequency), 0, Frequency))

ggplot(actor_director_filtered, aes(x = director, y = Actor, fill = Frequency)) +
  geom_tile() +
  
  theme_minimal() +
  
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 6),
    axis.text.y = element_text(size = 6)
  ) +
  
  scale_fill_gradientn(
  colors = c("white","grey","blue4", "green3", "gold", "red"),
  values = scales::rescale(c(0, 1, 2, 3, 4, 5))
) +
  geom_tile(color = "grey90")+
  
  labs(
    title = "Actor–Director Heatmap (Actors with ≥2 Appearances)",
    x = "Directors",
    y = "Actors",
    fill = "Frequency"
  )

```

#Chi-Squared Tests
#Keep only Winner and Nominee rows (all)
```{r}
award_data_all <- combined %>%
  filter(Award_Status %in% c("Winner", "Nominee"))
```

#Keep only Winner and Nominee rows (top 20)
```{r}
award_data <- combined_top %>%
  filter(Award_Status %in% c("Winner", "Nominee"))
```

# Create Actor x Director contingency table (all)
```{r}
actor_director_alltable <- table(
  award_data_all$Actor,
  award_data_all$director
)
```

# Create Actor x Director contingency table (top 20)
```{r}
actor_director_table <- table(
  award_data$Actor,
  award_data$director
)
```
# Run Chi-Square Test of Independence (all)
```{r}
chi_result_all <- chisq.test(actor_director_alltable)
```

# Run Chi-Square Test of Independence (top 20)
```{r}
chi_result <- chisq.test(actor_director_table)
```
# Standardized residuals (all)
```{r}
chi_result_all$residuals
```

# Standardized residuals (top 20)
```{r}
chi_result$residuals
```
# Convert standardized residuals into a dataframe (all)
```{r}
residual_table_all <- as.data.frame(as.table(chi_result_all$residuals))
```

# Convert standardized residuals into a dataframe (top 20)
```{r}
residual_table <- as.data.frame(as.table(chi_result$residuals))
```
# Rename Columns (all)
```{r}
colnames(residual_table_all) <- c("Actor", "Director", "Residual")
```
# Rename Columns (top 20)
```{r}
colnames(residual_table) <- c("Actor", "Director", "Residual")
```
# Sort by largest absolute residuals (all)
```{r}
residual_table_all <- residual_table_all %>%
  arrange(desc(abs(Residual)))
```
# Sort by largest absolute residuals (top 20)
```{r}
residual_table <- residual_table %>%
  arrange(desc(abs(Residual)))
```
## Monte Carlo simulated Chi-Square test (all)
```{r}
actor_director_table_all <- table(
  award_data_all$Actor,
  award_data_all$director
)

set.seed(123)

chi_result_sim_all <- chisq.test(
  actor_director_table_all,
  simulate.p.value = TRUE,
  B = 10000
)

chi_result_sim_all
```

## Monte Carlo simulated Chi-Square test (top 20)
```{r}
actor_director_table <- table(
  award_data$Actor,
  award_data$director
)

set.seed(123)  

chi_result_sim <- chisq.test(
  actor_director_table,
  simulate.p.value = TRUE,
  B = 10000
)

chi_result_sim
```
#Win probability with a Director (all)
```{r}
director_win_prob_all <- award_data_all %>%
  group_by(director) %>%
  summarise(
    total_collaborations = n(),
    wins = sum(Award_Status == "Winner"),
    win_probability = wins / total_collaborations
  ) %>%
  arrange(desc(win_probability))

director_win_prob_all
```
#Win probability Director (top 20)
```{r}
director_win_prob <- award_data %>%
  group_by(director) %>%
  summarise(
    total_collaborations = n(),
    wins = sum(Award_Status == "Winner"),
    win_probability = wins / total_collaborations
  ) %>%
  arrange(desc(win_probability))

director_win_prob
```
#Actor & Director Win Probability (all)
```{r}
actor_director_win_prob_all <- award_data_all %>%
  group_by(director, Actor) %>%
  summarise(
    total = n(),
    wins = sum(Award_Status == "Winner"),
    win_prob = wins / total,
    .groups = "drop"
  ) %>%
  arrange(desc(win_prob))

actor_director_win_prob_all
```
#Actor & Director Win Probability (top 20)
```{r}
actor_director_win_prob <- award_data %>%
  group_by(director, Actor) %>%
  summarise(
    total = n(),
    wins = sum(Award_Status == "Winner"),
    win_prob = wins / total,
    .groups = "drop"
  ) %>%
  arrange(desc(win_prob))

actor_director_win_prob
```
#Filtered by total >=3 (all)
```{r}
actor_director_win_prob_filtered_all <- actor_director_win_prob_all %>%
  filter(total >= 3)
```
#Filtered by total >=3 (top 20)
```{r}
actor_director_win_prob_filtered <- actor_director_win_prob %>%
  filter(total >= 3)
```

##Network Analysis
# All Director/Actor Edges
```{r}
edges_director_actor_all <- combined %>%
  select(Film, director, Actor) %>%
  distinct() %>%
  rename(from = director, to = Actor)
```
# Top 20 Director/Actor Edges
```{r}
edges_director_actor <- combined_top %>%
  select(Film, director, Actor) %>%
  distinct() %>%
  rename(from = director, to = Actor)
```
#All Director/Actor Nodes
```{r}
award_scores_all <- combined %>%
  group_by(director) %>%
  summarise(awards = sum(Status == "Winner", na.rm = TRUE)) %>%
  rename(name = director)

actor_scores_all <- combined %>%
  group_by(Actor) %>%
  summarise(awards = sum(Status == "Winner", na.rm = TRUE)) %>%
  rename(name = Actor)

nodes_all <- bind_rows(award_scores_all, actor_scores_all) %>%
  group_by(name) %>%
  summarise(awards = sum(awards, na.rm = TRUE))
```
#Top 20 Director/Actor Nodes
```{r}
award_scores <- combined_top %>%
  group_by(director) %>%
  summarise(awards = sum(Status == "Winner", na.rm = TRUE)) %>%
  rename(name = director)

actor_scores <- combined_top %>%
  group_by(Actor) %>%
  summarise(awards = sum(Status == "Winner", na.rm = TRUE)) %>%
  rename(name = Actor)

nodes <- bind_rows(award_scores, actor_scores) %>%
  group_by(name) %>%
  summarise(awards = sum(awards, na.rm = TRUE))
```

#Status to weighted score (all)
```{r}
na_all <- combined %>%
  mutate(
    Status = trimws(Status),
    award_score_all = case_when(
      Status == "Winner" ~ 2,
      Status == "Nominee" ~ 1,
      Status == "Non-Nominee" ~ 0,
      TRUE ~ 0
    )
  )
```
#Status to weighted score (top 20)
```{r}
na <- combined_top %>%
  mutate(
    Status = trimws(Status),
    award_score = case_when(
      Status == "Winner" ~ 2,
      Status == "Nominee" ~ 1,
      Status == "Non-Nominee" ~ 0,
      TRUE ~ 0
    )
  )
```
#Rebuild node scores (all)
```{r}
director_scores_all <- na_all %>%
  group_by(director) %>%
  summarise(awards = sum(award_score_all, na.rm = TRUE)) %>%
  rename(name = director)

actor_scores_all <- na_all %>%
  group_by(Actor) %>%
  summarise(awards = sum(award_score_all, na.rm = TRUE)) %>%
  rename(name = Actor)

node_attributes_all <- bind_rows(director_scores_all, actor_scores_all) %>%
  group_by(name) %>%
  summarise(awards = sum(awards), .groups = "drop")
```
#Rebuild node scores (top 20)
```{r}
director_scores <- na %>%
  group_by(director) %>%
  summarise(awards = sum(award_score, na.rm = TRUE)) %>%
  rename(name = director)

actor_scores <- na %>%
  group_by(Actor) %>%
  summarise(awards = sum(award_score, na.rm = TRUE)) %>%
  rename(name = Actor)

node_attributes <- bind_rows(director_scores, actor_scores) %>%
  group_by(name) %>%
  summarise(awards = sum(awards), .groups = "drop")
```
#Vertices from edges only (all)
```{r}
nodes_all <- data.frame(
  name = unique(c(edges_director_actor_all$from,
                  edges_director_actor_all$to))
)

nodes_all <- nodes_all %>%
  left_join(node_attributes_all, by = "name") %>%
  mutate(awards = ifelse(is.na(awards), 0, awards))
```
#Vertices from edges only (top 20)
```{r}
nodes <- data.frame(
  name = unique(c(edges_director_actor$from,
                  edges_director_actor$to))
)

nodes <- nodes %>%
  left_join(node_attributes, by = "name") %>%
  mutate(awards = ifelse(is.na(awards), 0, awards))
```

#Check (all)
```{r}
setdiff(edges_director_actor_all$from, nodes_all$name)
setdiff(edges_director_actor_all$to, nodes_all$name)
```
#Check(top 20)
```{r}
setdiff(edges_director_actor$from, nodes$name)
setdiff(edges_director_actor$to, nodes$name)
```
#Clean Names again as missing actor/directors in nodes (all)
```{r}
clean_na_all <- na_all %>%
  mutate(
    director = trimws(tolower(director)),
    Actor = trimws(tolower(Actor)),
    Status = trimws(Status)
  )
```
# Clean Names again as missing actor/directors in nodes (top 20)
```{r}
clean_na <- na %>%
  mutate(
    director = trimws(tolower(director)),
    Actor = trimws(tolower(Actor)),
    Status = trimws(Status)
  )
```

#Rebuild nodes from edges (all)
```{r}
edges_director_actor_all <- clean_na_all %>%
  select(director, Actor) %>%
  filter(!is.na(director), !is.na(Actor)) %>%
  filter(director != "", Actor != "") %>%
  distinct() %>%
  rename(from = director, to = Actor)
```
#Rebuild nodes from edges (top 20)
```{r}
edges_director_actor <- clean_na %>%
  select(director, Actor) %>%
  filter(!is.na(director), !is.na(Actor)) %>%
  filter(director != "", Actor != "") %>%
  distinct() %>%
  rename(from = director, to = Actor)
```
#Check (all)
```{r}
all(edges_director_actor_all$from %in% nodes_all$name)

all(edges_director_actor_all$to %in% nodes_all$name)
```
#Check (top 20)
```{r}
all(edges_director_actor$from %in% nodes$name)

all(edges_director_actor$to %in% nodes$name)
```
#Clean edge names again (all)
```{r}
edges_director_actor_all <- edges_director_actor_all %>%
  mutate(
    from = trimws(tolower(from)),
    to = trimws(tolower(to))
  ) %>%
  filter(
    from != "",
    to != "",
    !is.na(from),
    !is.na(to)
  )
```
# rebuild nodes from edges (all)
```{r}
nodes_all <- data.frame(
  name = unique(c(
    edges_director_actor_all$from,
    edges_director_actor_all$to
  ))
)
```
#check (all)
```{r}
all(edges_director_actor_all$from %in% nodes_all$name)

all(edges_director_actor_all$to %in% nodes_all$name)
```
#Clean edge names again (top 20)
```{r}
edges_director_actor <- edges_director_actor %>%
  mutate(
    from = trimws(tolower(from)),
    to = trimws(tolower(to))
  ) %>%
  filter(
    from != "",
    to != "",
    !is.na(from),
    !is.na(to)
  )
```
# rebuild nodes from edges (top 20)
```{r}
nodes <- data.frame(
  name = unique(c(
    edges_director_actor$from,
    edges_director_actor$to
  ))
)
```
#check (top 20)
```{r}
all(edges_director_actor$from %in% nodes$name)

all(edges_director_actor$to %in% nodes$name)
```
#Build graph
```{r}
graph_all <- graph_from_data_frame(
  d = edges_director_actor_all,
  vertices = nodes_all,
  directed = FALSE
)

nodes_vis_all <- data.frame(
  id = nodes_all$name,
  label = nodes_all$name
)

nodes_vis_all$degree <- degree(graph_all)

nodes_vis_all$value <- rescale(
  exp(nodes_vis_all$degree / max(nodes_vis_all$degree)),
  to = c(5, 40)
)

directors <- unique(edges_director_actor_all$from)

director_colors <- rainbow(length(directors))
names(director_colors) <- directors

nodes_vis_all$color <- ifelse(
  nodes_vis_all$id %in% directors,
  director_colors[nodes_vis_all$id],
  "lightblue"
  )

edges_vis_all <- edges_director_actor_all %>%
  group_by(from, to) %>%
  summarise(weight = n(), .groups = "drop")

edges_vis_all$value <- edges_vis_all$weight

visNetwork(nodes_vis_all, edges_vis_all) %>%
  
  visNodes(
    shape = "dot"
  ) %>%
  
  visEdges(
    smooth = FALSE,
    scaling = list(min = 1, max = 10),
    color = "gray"
  ) %>%
  
  visOptions(
    highlightNearest = TRUE,
    nodesIdSelection = TRUE
  ) %>%
  
  visPhysics(
    stabilization = TRUE
  )
```



