---
title: "AcademyAwards"
author: "Emely P."
date: "2026-05-16"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
#import library
```{r}
library(ggplot2)
library(dplyr)
library(stringr)
library(readr)
library(tidyr)
library(reshape2)
library(jsonlite)
library(igraph)
library(tidygraph)
library(ggraph)
library(visNetwork)
library(scales)
```


# import csv
```{r}
data <- read_csv("~/Downloads/film_details_v3.csv")
```
#Frequency Count of every unique value in column "director" 
```{r}
table(data$director)
```
#Sort by most frequent
```{r}
sort(table(data$director), decreasing = TRUE)
```
#Plot in bar graph
```{r}
barplot(table(data$director))
```
#plot directors with more than 1 nomination
```{r}
dir_counts <- sort(table(data$director), decreasing = TRUE)
dir_counts_filtered <- dir_counts[dir_counts > 1]

barplot(dir_counts_filtered,
        las = 2,
        col = "navy",
        main = "Frequent Directors",
        ylab = "Number of Movies")
```
#Change orientation, names on y-axis
```{r}
barplot(dir_counts_filtered,
        horiz = TRUE,
        col = "navy",
        main = "Director Nominations Count",
        xlab = "Number of Movies",
        las = 1,
        space = 0.5)
```
#change bar graph so directors names are visable
```{r}
max_chars <- max(nchar(names(dir_counts_filtered)))
par(mar = c(5, max_chars * 0.7, 4, 2))

barplot(dir_counts_filtered,
        horiz = TRUE,
        col = "navy",
        main = "Director's Film Nomination Count",
        xlab = "Number of Movies",
        las = 1,
        space = 0.5)
```
# Total number of directors with more than 1 nomination
```{r}
dir_counts <- table(data$director)
sum(dir_counts > 1)
```
#Table of director counts
```{r}
dir_counts <- table(data$director)
table(dir_counts)
```
#Plot of director counts
```{r}
barplot(table(dir_counts),
        col = "navy",
        main = "Director Frequency Distribution",
        xlab = "Number of Nominations",
        ylab = "Number of Directors")
```
#Plot of director counts with y-axis going to 200
```{r}
freq <- table(dir_counts)

bp <- barplot(freq,
              col = "navy",
              main = "Director Frequency Distribution",
              xlab = "Number of Nominations",
              ylab = "Number of Directors",
              ylim = c(0, 200))   # force y-axis to 200
```
#add counts on top of bar
```{r}
bp <- barplot(freq,
              col = "navy",
              main = "Director Frequency Distribution",
              xlab = "Number of Nominations",
              ylab = "Number of Directors",
              ylim = c(0, 200))   # force y-axis to 200
text(x = bp,
     y = freq,
     labels = freq,
     pos = 3,        
     cex = 0.8)
```
#Cast frequency
```{r}
all_cast <- unlist(data[paste0("Cast ", 1:15)])
all_cast <- na.omit(all_cast)
cast_counts <- table(all_cast)
cast_counts
```
#Sort cast by frequency
```{r}
sort(cast_counts, decreasing = TRUE)
```
#Cast frequency table
```{r}
cast_counts <- table(all_cast)
cast_counts_sorted <- sort(cast_counts, decreasing = TRUE)
cast_counts_sorted[cast_counts_sorted > 1]
```
#Plot cast distribution
```{r}
cast_freq_dist <- table(cast_counts)
cast_freq_dist
barplot(cast_freq_dist,
        col = "darkgreen",
        main = "Distribution of Actor Frequencies",
        xlab = "Number of Appearances",
        ylab = "Number of Actors")
```
#Add cast counts on of bars
```{r}
cast_freq_dist <- table(cast_counts)
bp2 <- barplot(cast_freq_dist,
              col = "darkgreen",
              main = "Distribution of Actor Frequencies",
              xlab = "Number of Appearances",
              ylab = "Number of Actors",
              ylim = c(0, 4000))
text(x = bp2,
     y = cast_freq_dist,
     labels = cast_freq_dist,
     pos = 3,      # above bars
     cex = 0.8)
```
#Nominated Actor/Actress Counts from Nominee Column
```{r}
AA_final <- read_csv("~/data/628_AcademyAwards/AA_final.csv")
nominee_counts <- table(AA_final$Nominee)
nominee_counts
```
#Nominee counts for nominations only of Actors/Actress in Nominee column
```{r}
data_role <- AA_final[grepl("role", AA_final$Category, ignore.case = TRUE), ]
nominee_counts <- table(data_role$Nominee)
nominee_counts
```
#Nominee counts for nomination only of Actors/Actress in Nominee Column
```{r}
data_role <- AA_final[grepl("role", AA_final$Category, ignore.case = TRUE), ]
nominee_counts <- table(data_role$Nominee)
nominee_counts
```


#Remove actress in leading role entry, and sort
```{r}
nominee_counts <- table(AA_final$Nominee)
nominee_counts <- nominee_counts[!grepl("ACTRESS IN A LEADING ROLE",
                                        names(nominee_counts),
                                        ignore.case = TRUE)]
sort(nominee_counts, decreasing = TRUE)
```
#Count nominee frequency bar chart
```{r}
barplot(nominee_counts,
        horiz = TRUE,
        col = "pink",
        main = "Nominee Frequency",
        xlab = "Number of Nominations",
        las = 1)
```
#Nominee frequency distribution
```{r}
freq_dist <- table(nominee_counts)
freq_dist
barplot(freq_dist,
        col = "pink",
        xlab = "Number of Actor/Actress Nominations",
        ylab = "Number of Nominees",
        main = "Nominee Frequency Distribution")
```
#Add labels to bar chart
```{r}
bp3 <- barplot(freq_dist,
              col = "pink",
              ylim = c(0, max(freq_dist) * 1.1),
              xlab = "Number of Nominations",
              ylab = "Number of Nominees",
              main = "Nominee Frequency Distribution")

text(bp3,
     freq_dist,
     labels = freq_dist,
     pos = 3,
     cex = 0.8)
```
#Winner Frequency
```{r}
winners_role <- AA_final[
  AA_final$Award_Status == "Winner" &
  grepl("ROLE", AA_final$Category, ignore.case = TRUE),
]
winner_role_counts <- table(winners_role$Nominee)
sort(winner_role_counts, decreasing = TRUE)
```
#Plot Winner Frequency
```{r}
win_distribution <- table(winner_role_counts)
win_distribution
barplot(win_distribution,
        col = "chocolate",
        main = "Win Distribution",
        xlab = "Number of Wins",
        ylab = "Number of Winners")
```
#Add labels to winner frequency
```{r}
bp4 <- barplot(win_distribution,
              col = "chocolate",
              ylim = c(0, max(win_distribution) * 1.1),
              main = "Win Distribution",
              xlab = "Number of Wins",
              ylab = "Number of Winners")

text(bp4,
     win_distribution,
     labels = win_distribution,
     pos = 3,
     cex = 0.8)
```
#Directors with 4 or more associated nominations
```{r}
director_counts <- table(data$director)
topdirectors <- director_counts[director_counts >= 4]
topdirectors
```
#table of top directors
```{r}
data.frame(
  Director = names(topdirectors),
  Films = as.numeric(topdirectors)
)
```
#Filter top direcs in data.csv
```{r}
top_directors <- unique(data$director)
data_top <- data[data$director %in% top_directors, ]
```
#Verify count
```{r}
nrow(data_top)
unique(data_top$director)
```
#Cast List
```{r}
all_cast <- unlist(data[paste0("Cast ", 1:15)])
all_cast <- trimws(na.omit(all_cast))
all_cast <- all_cast[all_cast != ""]
unique_cast <- unique(all_cast)
```
#Split AA_Final Status
```{r}
all_people <- unique(AA_final$Award_Status)
winners <- unique(AA_final$Nominee[AA_final$Award_Status == "Winner"])
nominees <- unique(AA_final$Nominee[AA_final$Award_Status == "Nominee"])
```
#Cast not nominated
```{r}
all_award_people <- unique(AA_final$Award_Status)
not_nominated <- setdiff(unique_cast, all_award_people)
```
#Final Counts
```{r}
data.frame(
  Category = c("Won", "Nominated", "Not nominated"),
  Count = c(
    length(winners),
    length(nominees),
    length(not_nominated)
  )
)
```
# Change data.csv title column to film
```{r}
names(data)[names(data) == "title"] <- "Film"
```
#film-details import, change title to Film, and reformat
```{r}
film_details <- read_csv("~/Downloads/film_details.csv")
names(film_details)[names(film_details) == "title"] <- "Film"
cast <- film_details %>%
  separate_rows(top_cast, sep = ",") %>%
  mutate(Actor = trimws(top_cast)) %>%
  select(Film, director, Actor)
```
#Join cast & data role 
```{r}
combined <- cast %>%
  left_join(data_role, by = c("Film", "Actor" = "Nominee"))
```
#Stacked categories, Winner, Nominee, Non-Nominee
```{r}
combined <- combined %>%
  mutate(Status = case_when(
    Award_Status == "Winner" ~ "Winner",
    Award_Status == "Nominee" ~ "Nominee",
    TRUE ~ "Non-Nominee"
  ))
```
#plot stacked bar chart
```{r}
ggplot(combined, aes(x = director, fill = Status)) +
  geom_bar() +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
ggsave("plot2.png", width = 14, height = 8, dpi = 300)
```
#Switch x & y axis
```{r}
ggplot(combined, aes(x = director, fill = Status)) +
  geom_bar() +
  
  coord_flip() +
  
  theme_minimal() +
  
  theme(
    axis.text.y = element_text(size = 6),
    plot.margin = margin(10, 20, 20, 20)
  )

ggsave("plot2.png", width = 14, height = 14, dpi = 300)
```
#Filter director list
```{r}
film_filtered <- cast %>%
  filter(director %in% data$director)
```
#combined award data
```{r}
combined <- film_filtered %>%
  left_join(data_role, by = c("Film", "Actor" = "Nominee"))
```

#category list
```{r}
combined <- combined %>%
  mutate(Status = case_when(
    Award_Status == "Winner" ~ "Winner",
    Award_Status == "Nominee" ~ "Nominee",
    TRUE ~ "Non-Nominee"
  ))
```
#Plot stacked bar chart
```{r}
ggplot(combined, aes(x = director, fill = Status)) +
  geom_bar() +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

#top 20 directors
```{r}
top_directors <- combined %>%
  count(director, sort = TRUE) %>%
  slice_head(n = 20) %>%
  pull(director)

combined_top <- combined %>%
  filter(director %in% top_directors)

ggplot(combined_top, aes(x = reorder(director, -table(director)[director]), 
                         fill = Status)) +
  geom_bar() +
  theme_minimal() +
  labs(x = "Director", y = "Count") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
#Change order status
```{r}
combined$Status <- factor(
  combined$Status,
  levels = c("Winner","Nominee","Non-Nominee")
)
```
#Plot stacked bar chart
```{r}
top_directors <- combined %>%
  count(director, sort = TRUE) %>%
  slice_head(n = 20) %>%
  pull(director)

combined_top <- combined %>%
  filter(director %in% top_directors)

ggplot(combined_top, aes(x = reorder(director, -table(director)[director]), 
                         fill = Status)) +
  geom_bar() +
  theme_minimal() +
  labs(x = "Director", y = "Count") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

#Percentage stacked bar chart
```{r}
ggplot(combined_top, aes(x = director, fill = Status)) +
  geom_bar(position = "fill") +
  scale_y_continuous(labels = scales::percent) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
#count stacked bar chart 
```{r}
ggplot(combined_top, aes(x = director, fill = Status)) +
  geom_bar() +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(y = "Count", x = "Director")
```
summary table
```{r}
label_data <- combined_top %>%
  group_by(director, Status) %>%
  summarise(n = n(), .groups = "drop") %>%
  group_by(director) %>%
  mutate(total = sum(n))
```
#labels
```{r}
label_data <- label_data %>%
  mutate(label = n)
```
#stacked bar chart with labels
```{r}
ggplot(label_data, aes(x = director, y = n, fill = Status)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = label),
            position = position_stack(vjust = 0.5),
            size = 2) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(y = "Count", x = "Director")
```
#Director film count
```{r}
director_film_count_all <- combined %>%
  group_by(director) %>%
  summarise(film_count = n_distinct(Film))
print(director_film_count)
```
#top 20 director film count
```{r}
director_film_count <- combined_top %>%
  group_by(director) %>%
  summarise(film_count = n_distinct(Film))
print(director_film_count)
```
#All Actor/Actress frequency
```{r}
actor_frequency_all <- combined %>%
  group_by(Actor) %>%
  summarise(Frequency = n(), .groups = "drop") %>%
  arrange(desc(Frequency))
print(actor_frequency_all, n = Inf)
```

# Actor/Actress frequency in top 20 directors
```{r}
actor_frequency <- combined_top %>%
  group_by(Actor) %>%
  summarise(Frequency = n(), .groups = "drop") %>%
  arrange(desc(Frequency))
print(actor_frequency, n = Inf)
```
# Heat map code
#All Actor-Director Counts
```{r}
actor_director_all <- combined %>%
  group_by(director, Actor) %>%
  summarise(Frequency = n(), .groups = "drop")
```

#Top 20 Actor-Directo Counts
```{r}
actor_director <- combined_top %>%
  group_by(director, Actor) %>%
  summarise(Frequency = n(), .groups = "drop")
```

#Plot Heat Map All 
```{r}
ggplot(actor_director_all, aes(x = director, y = Actor, fill = Frequency)) +
  geom_tile() +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 6),
    axis.text.y = element_text(size = 6)
  ) +
  scale_fill_gradient(low = "white", high = "steelblue") +
  labs(title = "Actor–Director Collaboration Heatmap",
       x = "Director",
       y = "Actor",
       fill = "Frequency")
```
#Heat map of top 20 
```{r}
ggplot(actor_director, aes(x = director, y = Actor, fill = Frequency)) +
  geom_tile() +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 6),
    axis.text.y = element_text(size = 6)
  ) +
  scale_fill_gradient(low = "white", high = "steelblue") +
  labs(title = "Actor–Director Collaboration Heatmap",
       x = "Director",
       y = "Actor",
       fill = "Frequency")
```
# Heat Map with Actors with 2 or more counts (in all actors)
```{r}
valid_actors_all <- combined %>%
  count(Actor) %>%
  filter(n >= 2) %>%
  pull(Actor)

actor_director_filtered_all <- combined %>%
  filter(Actor %in% valid_actors_all) %>%
  count(director, Actor, name = "Frequency")

full_grid_all <- expand.grid(
  director = unique(actor_director_filtered_all$director),
  Actor = unique(actor_director_filtered_all$Actor)
)

actor_director_filtered_all <- full_grid_all %>%
  left_join(actor_director_filtered_all,
            by = c("director", "Actor")) %>%
  mutate(Frequency = ifelse(is.na(Frequency), 0, Frequency))

ggplot(actor_director_filtered_all, aes(x = director, y = Actor, fill = Frequency)) +
  geom_tile() +
  
  theme_minimal() +
  
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 6),
    axis.text.y = element_text(size = 6)
  ) +
  
  scale_fill_gradientn(
  colors = c("white","grey","blue4", "green3", "gold", "red"),
  values = scales::rescale(c(0, 1, 2, 3, 4, 5))
) +
  geom_tile(color = "grey90")+
  
  labs(
    title = "All Actor–Director Heatmap (Actors with ≥2 Appearances)",
    x = "Directors",
    y = "Actors",
    fill = "Frequency"
  )
```
# Heat map with actors with 4 or more counts in top 20 directors
```{r}
valid_actors <- combined_top %>%
  count(Actor) %>%
  filter(n >= 4) %>%
  pull(Actor)

actor_director_filtered <- combined_top %>%
  filter(Actor %in% valid_actors) %>%
  count(director, Actor, name = "Frequency")

full_grid <- expand.grid(
  director = unique(actor_director_filtered$director),
  Actor = unique(actor_director_filtered$Actor)
)

actor_director_filtered <- full_grid %>%
  left_join(actor_director_filtered,
            by = c("director", "Actor")) %>%
  mutate(Frequency = ifelse(is.na(Frequency), 0, Frequency))

ggplot(actor_director_filtered, aes(x = director, y = Actor, fill = Frequency)) +
  geom_tile() +
  
  theme_minimal() +
  
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 6),
    axis.text.y = element_text(size = 6)
  ) +
  
  scale_fill_gradientn(
  colors = c("white","grey","blue4", "green3", "gold", "red"),
  values = scales::rescale(c(0, 1, 2, 3, 4, 5))
) +
  geom_tile(color = "grey90")+
  
  labs(
    title = "Actor–Director Heatmap (Actors with ≥2 Appearances)",
    x = "Directors",
    y = "Actors",
    fill = "Frequency"
  )

```

#Chi-Squared Tests
#Keep only Winner and Nominee rows (all)
```{r}
award_data_all <- combined %>%
  filter(Award_Status %in% c("Winner", "Nominee"))
```

#Keep only Winner and Nominee rows (top 20)
```{r}
award_data <- combined_top %>%
  filter(Award_Status %in% c("Winner", "Nominee"))
```

# Create Actor x Director contingency table (all)
```{r}
actor_director_alltable <- table(
  award_data_all$Actor,
  award_data_all$director
)
```

# Create Actor x Director contingency table (top 20)
```{r}
actor_director_table <- table(
  award_data$Actor,
  award_data$director
)
```
# Run Chi-Square Test of Independence (all)
```{r}
chi_result_all <- chisq.test(actor_director_alltable)
```

# Run Chi-Square Test of Independence (top 20)
```{r}
chi_result <- chisq.test(actor_director_table)
```
# Standardized residuals (all)
```{r}
chi_result_all$residuals
```

# Standardized residuals (top 20)
```{r}
chi_result$residuals
```
# Convert standardized residuals into a dataframe (all)
```{r}
residual_table_all <- as.data.frame(as.table(chi_result_all$residuals))
```

# Convert standardized residuals into a dataframe (top 20)
```{r}
residual_table <- as.data.frame(as.table(chi_result$residuals))
```
# Rename Columns (all)
```{r}
colnames(residual_table_all) <- c("Actor", "Director", "Residual")
```
# Rename Columns (top 20)
```{r}
colnames(residual_table) <- c("Actor", "Director", "Residual")
```
# Sort by largest absolute residuals (all)
```{r}
residual_table_all <- residual_table_all %>%
  arrange(desc(abs(Residual)))
```
# Sort by largest absolute residuals (top 20)
```{r}
residual_table <- residual_table %>%
  arrange(desc(abs(Residual)))
```
## Monte Carlo simulated Chi-Square test (all)
```{r}
actor_director_table_all <- table(
  award_data_all$Actor,
  award_data_all$director
)

set.seed(123)

chi_result_sim_all <- chisq.test(
  actor_director_table_all,
  simulate.p.value = TRUE,
  B = 10000
)

chi_result_sim_all
```

## Monte Carlo simulated Chi-Square test (top 20)
```{r}
actor_director_table <- table(
  award_data$Actor,
  award_data$director
)

set.seed(123)  

chi_result_sim <- chisq.test(
  actor_director_table,
  simulate.p.value = TRUE,
  B = 10000
)

chi_result_sim
```
#Win probability with a Director (all)
```{r}
director_win_prob_all <- award_data_all %>%
  group_by(director) %>%
  summarise(
    total_collaborations = n(),
    wins = sum(Award_Status == "Winner"),
    win_probability = wins / total_collaborations
  ) %>%
  arrange(desc(win_probability))

director_win_prob_all
```
#Win probability Director (top 20)
```{r}
director_win_prob <- award_data %>%
  group_by(director) %>%
  summarise(
    total_collaborations = n(),
    wins = sum(Award_Status == "Winner"),
    win_probability = wins / total_collaborations
  ) %>%
  arrange(desc(win_probability))

director_win_prob
```
#Actor & Director Win Probability (all)
```{r}
actor_director_win_prob_all <- award_data_all %>%
  group_by(director, Actor) %>%
  summarise(
    total = n(),
    wins = sum(Award_Status == "Winner"),
    win_prob = wins / total,
    .groups = "drop"
  ) %>%
  arrange(desc(win_prob))

actor_director_win_prob_all
```
#Actor & Director Win Probability (top 20)
```{r}
actor_director_win_prob <- award_data %>%
  group_by(director, Actor) %>%
  summarise(
    total = n(),
    wins = sum(Award_Status == "Winner"),
    win_prob = wins / total,
    .groups = "drop"
  ) %>%
  arrange(desc(win_prob))

actor_director_win_prob
```
#Filtered by total >=3 (all)
```{r}
actor_director_win_prob_filtered_all <- actor_director_win_prob_all %>%
  filter(total >= 3)
```
#Filtered by total >=3 (top 20)
```{r}
actor_director_win_prob_filtered <- actor_director_win_prob %>%
  filter(total >= 3)
```

##Network Analysis
# All Director/Actor Edges
```{r}
edges_director_actor_all <- combined %>%
  select(Film, director, Actor) %>%
  distinct() %>%
  rename(from = director, to = Actor)
```
# Top 20 Director/Actor Edges
```{r}
edges_director_actor <- combined_top %>%
  select(Film, director, Actor) %>%
  distinct() %>%
  rename(from = director, to = Actor)
```
#All Director/Actor Nodes
```{r}
award_scores_all <- combined %>%
  group_by(director) %>%
  summarise(awards = sum(Status == "Winner", na.rm = TRUE)) %>%
  rename(name = director)

actor_scores_all <- combined %>%
  group_by(Actor) %>%
  summarise(awards = sum(Status == "Winner", na.rm = TRUE)) %>%
  rename(name = Actor)

nodes_all <- bind_rows(award_scores_all, actor_scores_all) %>%
  group_by(name) %>%
  summarise(awards = sum(awards, na.rm = TRUE))
```
#Top 20 Director/Actor Nodes
```{r}
award_scores <- combined_top %>%
  group_by(director) %>%
  summarise(awards = sum(Status == "Winner", na.rm = TRUE)) %>%
  rename(name = director)

actor_scores <- combined_top %>%
  group_by(Actor) %>%
  summarise(awards = sum(Status == "Winner", na.rm = TRUE)) %>%
  rename(name = Actor)

nodes <- bind_rows(award_scores, actor_scores) %>%
  group_by(name) %>%
  summarise(awards = sum(awards, na.rm = TRUE))
```

#Status to weighted score (all)
```{r}
na_all <- combined %>%
  mutate(
    Status = trimws(Status),
    award_score_all = case_when(
      Status == "Winner" ~ 2,
      Status == "Nominee" ~ 1,
      Status == "Non-Nominee" ~ 0,
      TRUE ~ 0
    )
  )
```
#Status to weighted score (top 20)
```{r}
na <- combined_top %>%
  mutate(
    Status = trimws(Status),
    award_score = case_when(
      Status == "Winner" ~ 2,
      Status == "Nominee" ~ 1,
      Status == "Non-Nominee" ~ 0,
      TRUE ~ 0
    )
  )
```
#Rebuild node scores (all)
```{r}
director_scores_all <- na_all %>%
  group_by(director) %>%
  summarise(awards = sum(award_score_all, na.rm = TRUE)) %>%
  rename(name = director)

actor_scores_all <- na_all %>%
  group_by(Actor) %>%
  summarise(awards = sum(award_score_all, na.rm = TRUE)) %>%
  rename(name = Actor)

node_attributes_all <- bind_rows(director_scores_all, actor_scores_all) %>%
  group_by(name) %>%
  summarise(awards = sum(awards), .groups = "drop")
```
#Rebuild node scores (top 20)
```{r}
director_scores <- na %>%
  group_by(director) %>%
  summarise(awards = sum(award_score, na.rm = TRUE)) %>%
  rename(name = director)

actor_scores <- na %>%
  group_by(Actor) %>%
  summarise(awards = sum(award_score, na.rm = TRUE)) %>%
  rename(name = Actor)

node_attributes <- bind_rows(director_scores, actor_scores) %>%
  group_by(name) %>%
  summarise(awards = sum(awards), .groups = "drop")
```
#Vertices from edges only (all)
```{r}
nodes_all <- data.frame(
  name = unique(c(edges_director_actor_all$from,
                  edges_director_actor_all$to))
)

nodes_all <- nodes_all %>%
  left_join(node_attributes_all, by = "name") %>%
  mutate(awards = ifelse(is.na(awards), 0, awards))
```
#Vertices from edges only (top 20)
```{r}
nodes <- data.frame(
  name = unique(c(edges_director_actor$from,
                  edges_director_actor$to))
)

nodes <- nodes %>%
  left_join(node_attributes, by = "name") %>%
  mutate(awards = ifelse(is.na(awards), 0, awards))
```

#Check (all)
```{r}
setdiff(edges_director_actor_all$from, nodes_all$name)
setdiff(edges_director_actor_all$to, nodes_all$name)
```
#Check(top 20)
```{r}
setdiff(edges_director_actor$from, nodes$name)
setdiff(edges_director_actor$to, nodes$name)
```
#Clean Names again as missing actor/directors in nodes (all)
```{r}
clean_na_all <- na_all %>%
  mutate(
    director = trimws(tolower(director)),
    Actor = trimws(tolower(Actor)),
    Status = trimws(Status)
  )
```
# Clean Names again as missing actor/directors in nodes (top 20)
```{r}
clean_na <- na %>%
  mutate(
    director = trimws(tolower(director)),
    Actor = trimws(tolower(Actor)),
    Status = trimws(Status)
  )
```

#Rebuild nodes from edges (all)
```{r}
edges_director_actor_all <- clean_na_all %>%
  select(director, Actor) %>%
  filter(!is.na(director), !is.na(Actor)) %>%
  filter(director != "", Actor != "") %>%
  distinct() %>%
  rename(from = director, to = Actor)
```
#Rebuild nodes from edges (top 20)
```{r}
edges_director_actor <- clean_na %>%
  select(director, Actor) %>%
  filter(!is.na(director), !is.na(Actor)) %>%
  filter(director != "", Actor != "") %>%
  distinct() %>%
  rename(from = director, to = Actor)
```
#Check (all)
```{r}
all(edges_director_actor_all$from %in% nodes_all$name)

all(edges_director_actor_all$to %in% nodes_all$name)
```
#Check (top 20)
```{r}
all(edges_director_actor$from %in% nodes$name)

all(edges_director_actor$to %in% nodes$name)
```
#Clean edge names again (all)
```{r}
edges_director_actor_all <- edges_director_actor_all %>%
  mutate(
    from = trimws(tolower(from)),
    to = trimws(tolower(to))
  ) %>%
  filter(
    from != "",
    to != "",
    !is.na(from),
    !is.na(to)
  )
```
# rebuild nodes from edges (all)
```{r}
nodes_all <- data.frame(
  name = unique(c(
    edges_director_actor_all$from,
    edges_director_actor_all$to
  ))
)
```
#check (all)
```{r}
all(edges_director_actor_all$from %in% nodes_all$name)

all(edges_director_actor_all$to %in% nodes_all$name)
```
#Clean edge names again (top 20)
```{r}
edges_director_actor <- edges_director_actor %>%
  mutate(
    from = trimws(tolower(from)),
    to = trimws(tolower(to))
  ) %>%
  filter(
    from != "",
    to != "",
    !is.na(from),
    !is.na(to)
  )
```
# rebuild nodes from edges (top 20)
```{r}
nodes <- data.frame(
  name = unique(c(
    edges_director_actor$from,
    edges_director_actor$to
  ))
)
```
#check (top 20)
```{r}
all(edges_director_actor$from %in% nodes$name)

all(edges_director_actor$to %in% nodes$name)
```
#Build graph (all)
```{r}
graph_all <- graph_from_data_frame(
  d = edges_director_actor_all,
  vertices = nodes_all,
  directed = FALSE
)

nodes_vis_all <- data.frame(
  id = nodes_all$name,
  label = nodes_all$name
)

nodes_vis_all$degree <- degree(graph_all)

nodes_vis_all$value <- rescale(
  exp(nodes_vis_all$degree / max(nodes_vis_all$degree)),
  to = c(5, 40)
)

directors <- unique(edges_director_actor_all$from)

director_colors <- rainbow(length(directors))
names(director_colors) <- directors

nodes_vis_all$color <- ifelse(
  nodes_vis_all$id %in% directors,
  director_colors[nodes_vis_all$id],
  "lightblue"
  )

edges_vis_all <- edges_director_actor_all %>%
  group_by(from, to) %>%
  summarise(weight = n(), .groups = "drop")

edges_vis_all$value <- edges_vis_all$weight

visNetwork(nodes_vis_all, edges_vis_all) %>%
  
  visNodes(
    shape = "dot"
  ) %>%
  
  visEdges(
    smooth = FALSE,
    scaling = list(min = 1, max = 10),
    color = "gray"
  ) %>%
  
  visOptions(
    highlightNearest = TRUE,
    nodesIdSelection = TRUE
  ) %>%
  
  visPhysics(
    stabilization = TRUE
  )
```

#Build graph (top 20)
```{r}
graph <- graph_from_data_frame(
  d = edges_director_actor,
  vertices = nodes,
  directed = FALSE
)

nodes_vis <- data.frame(
  id = nodes$name,
  label = nodes$name
)

nodes_vis$degree <- degree(graph)

nodes_vis$value <- rescale(
  exp(nodes_vis$degree / max(nodes_vis$degree)),
  to = c(5, 40)
)

directors <- unique(edges_director_actor$from)

director_colors <- rainbow(length(directors))
names(director_colors) <- directors

nodes_vis$color <- ifelse(
  nodes_vis$id %in% directors,
  director_colors[nodes_vis$id],
  "lightblue"
  )

edges_vis <- edges_director_actor %>%
  group_by(from, to) %>%
  summarise(weight = n(), .groups = "drop")

edges_vis$value <- edges_vis$weight

visNetwork(nodes_vis, edges_vis) %>%
  
  visNodes(
    shape = "dot"
  ) %>%
  
  visEdges(
    smooth = FALSE,
    scaling = list(min = 1, max = 10),
    color = "gray"
  ) %>%
  
  visOptions(
    highlightNearest = TRUE,
    nodesIdSelection = TRUE
  ) %>%
  
  visPhysics(
    stabilization = TRUE
  )
```
#Chi-Square Analysis

#Director & Award outcome contingency table (all)
```{r}
combined$award <- ifelse(combined$Status %in% c("Winner", "Nominee"),
                    "Recognized", "Not_Recognized")
combined$award <- factor(combined$award, levels = c("Not_Recognized", "Recognized"))
```
#Director & Award outcome contingency table (top 20)
```{r}
combined_top$award <- ifelse(combined_top$Status %in% c("Winner", "Nominee"),
                    "Recognized", "Not_Recognized")
combined_top$award <- factor(combined_top$award, levels = c("Not_Recognized", "Recognized"))
```

#Generate contingency table (all)
```{r}
tab_all <- xtabs(~ director + award, data = combined)  
tab_all
```
#Generate contingency table (top 20)
```{r}
tab <- xtabs(~ director + award, data = combined_top)  
tab
```
#Add Row proportions: within each director, what % are recognized? (all)
```{r}
prop_by_director_all <- prop.table(tab_all, margin = 1)
prop_by_director_all
```
#Add Row proportions: within each director, what % are recognized? (top 20)
```{r}
prop_by_director <- prop.table(tab, margin = 1)
prop_by_director
```
#Chi-square test of independence (all)
```{r}
chisq.test(tab_all)
```

#Chi-square test of independence (top 20)
```{r}
chisq.test(tab)
```
#Check expected counts (all)
```{r}
test_all <- chisq.test(tab_all)
test_all$expected
```

#Check expected counts (top 20)
```{r}
test <- chisq.test(tab)
test$expected
```
#Subset of directors (top 10)
```{r}
top10_directors <- names(sort(table(combined$director), decreasing = TRUE))[1:10]
combined_top10 <- combined[combined$director %in% top10_directors, ]

tab_top10 <- xtabs(~ director + award, data = combined_top10)
tab_top10

chisq.test(tab_top10)
prop.table(tab_top10, 1)
```
#Subset of directors (top 15)
```{r}
top15_directors <- names(sort(table(combined$director), decreasing = TRUE))[1:15]
combined_top15 <- combined[combined$director %in% top15_directors, ]

tab_top15 <- xtabs(~ director + award, data = combined_top15)
tab_top15

chisq.test(tab_top15)
prop.table(tab_top15, 1)
```

#Fisher's exact test -Alejandro González Iñárritu
```{r}
target <- "Alejandro González Iñárritu"  

combined$dir2 <- ifelse(combined$director == target, target, "Other")

tab_2x2 <- table(combined$dir2, combined$award)
tab_2x2

fisher.test(tab_2x2)
```
## Explanation: Recognition rate = (9/75) 12%, Other recognition rate= 8.27%, Null hypothesis (H0): no association (odds ratio = 1), p = 0.2871 is greater than 0.05, so you fail to reject H0, estimated odds ratio is 0.66, but with a wide CI crossing 1, so the direction and size of any effect are uncertain.

#Fisher's exact test -Christopher Nolan
```{r}
target <- "Christopher Nolan"  

combined$dir2 <- ifelse(combined$director == target, target, "Other")

tab_2x2 <- table(combined$dir2, combined$award)
tab_2x2

fisher.test(tab_2x2)
```
#Fisher's exact test -Clint Eastwood
```{r}
target <- "Clint Eastwood"  

combined$dir2 <- ifelse(combined$director == target, target, "Other")

tab_2x2 <- table(combined$dir2, combined$award)
tab_2x2

fisher.test(tab_2x2)
```

#Fisher's exact test -David Fincher 
```{r}
target <- "David Fincher"  

combined$dir2 <- ifelse(combined$director == target, target, "Other")

tab_2x2 <- table(combined$dir2, combined$award)
tab_2x2

fisher.test(tab_2x2)
```

#Fisher's exact test -Martin Scorsese 
```{r}
target <- "Martin Scorsese"  

combined$dir2 <- ifelse(combined$director == target, target, "Other")

tab_2x2 <- table(combined$dir2, combined$award)
tab_2x2

fisher.test(tab_2x2)
```

#Fisher's exact test -Paul Thomas Anderson
```{r}
target <- "Paul Thomas Anderson"  

combined$dir2 <- ifelse(combined$director == target, target, "Other")

tab_2x2 <- table(combined$dir2, combined$award)
tab_2x2

fisher.test(tab_2x2)
```

#Fisher's exact test -Quentin Tarantino
```{r}
target <- "Quentin Tarantino"  

combined$dir2 <- ifelse(combined$director == target, target, "Other")

tab_2x2 <- table(combined$dir2, combined$award)
tab_2x2

fisher.test(tab_2x2)
```

#Fisher's exact test -Ridley Scott
```{r}
target <- "Ridley Scott"  

combined$dir2 <- ifelse(combined$director == target, target, "Other")

tab_2x2 <- table(combined$dir2, combined$award)
tab_2x2

fisher.test(tab_2x2)
```

#Fisher's exact test -Steven Spielberg
```{r}
target <- "Steven Spielberg"  

combined$dir2 <- ifelse(combined$director == target, target, "Other")

tab_2x2 <- table(combined$dir2, combined$award)
tab_2x2

fisher.test(tab_2x2)
```
#Recognition by top 20 director
```{r}
topN <- 20

top_directors <- names(sort(table(combined$director), decreasing = TRUE))[1:topN]
dat_top <- subset(combined, director %in% top_directors)

rate_top <- prop.table(tab, margin = 1)
rate_top

recognition_rate <- data.frame(
  director = rownames(tab),
  n_total = rowSums(tab),
  n_recognized = tab[, "Recognized"],
  recognized_rate = tab[, "Recognized"] / rowSums(tab)
)

recognition_rate[order(-recognition_rate$recognized_rate), ]
```
#Chi-Square test top 20
```{r}
chi <- chisq.test(tab)
chi
```
# Check expected counts (top 20)
```{r}
min_expected <- min(chi$expected)
prop_small <- mean(chi$expected < 5)

min_expected
prop_small
```
#Simulated p-value (top 20). The (simulated) p-value is significant (e.g., p < 0.05). If p-value is not significant: No evidence that recognition depends on director among the Top N.
```{r}
chi_sim <- chisq.test(tab, simulate.p.value = TRUE, B = 10000)
chi_sim
```
#Which directors contribute most to the association using standardized residuals. Large positive residuals: more recognized than expected for that director.Large negative residuals: fewer recognized than expected.
```{r}
chi_sim$stdres  
```
#Rank directors by residual (recognized)
```{r}
std <- chi_sim$stdres
sort(std[, "Recognized"], decreasing = TRUE)
```
#Residenuals and recognition rates
```{r}
tab_top <- tab  
rate <- tab_top[, "Recognized"] / rowSums(tab_top)

out <- data.frame(
  director = rownames(tab_top),
  n_total = rowSums(tab_top),
  recognized_rate = rate,
  stdres_recognized = chi_sim$stdres[, "Recognized"]
)

out[order(-out$stdres_recognized), ]
```
#Film recognition indicator, film have any nominee/winner in the cast
```{r}
combined$recognized <- combined$Status %in% c("Winner", "Nominee")

film_any_rec <- aggregate(recognized ~ Film + director, combined, any)
table(film_any_rec$recognized)
prop.table(table(film_any_rec$recognized))
```
# how many nominees/winners per film, more than 3
```{r}
film_n_rec <- aggregate(recognized ~ Film + director, combined, sum)
summary(film_n_rec$recognized)
head(film_n_rec[order(-film_n_rec$recognized), ], 40)
```
#films/row per actor
```{r}
actor_counts <- sort(table(combined$Actor), decreasing = TRUE)
head(actor_counts, 75)
```
#Counts of nominations/wins by category
```{r}
recognized_rows <- subset(combined, Status %in% c("Winner", "Nominee"))

sort(table(recognized_rows$Category), decreasing = TRUE)
```
#Wins vs nominees by category
```{r}
with(recognized_rows, table(Category, Status))
```
#Nominations/wins over time
```{r}
recognized_rows$Ceremony_Year <- as.integer(recognized_rows$Ceremony_Year)

sort(table(recognized_rows$Ceremony_Year), decreasing = FALSE)

with(recognized_rows, table(Ceremony_Year, Status))
```
#categories associated with directors
```{r}
tab_dir_cat <- with(recognized_rows, table(director, Category))

recognized_rows <- combined %>% 
  filter(Status %in% c("Winner", "Nominee"), !is.na(Category))

tab_dir_cat2 <- tab_dir_cat[rowSums(tab_dir_cat) > 0, colSums(tab_dir_cat) > 0]

print(tab_dir_cat2)
```
#Category proportion per director (all)
```{r}
dir_cat_profile <- prop.table(tab_dir_cat2, margin = 1)  


profile_df <- as.data.frame(dir_cat_profile)
names(profile_df) <- c("director", "Category", "prop")


counts_df <- as.data.frame(tab_dir_cat2)
names(counts_df) <- c("director", "Category", "n")

profile_df <- profile_df %>%
  left_join(counts_df, by = c("director", "Category")) %>%
  arrange(director, desc(prop))

profile_df
```
# Table for top 5 directors per category
```{r}
counts_df <- as.data.frame(tab_dir_cat2)
names(counts_df) <- c("director", "Category", "n")

top5_by_category_counts <- counts_df %>%
  group_by(Category) %>%
  arrange(desc(n), .by_group = TRUE) %>%
  slice_head(n = 5) %>%
  ungroup()

top5_by_category_counts
```
#actors had any prior nomination/win before this film?

#find film year for each film
```{r}
film_year <- combined %>%
  filter(!is.na(Ceremony_Year)) %>%
  group_by(Film, director) %>%
  summarise(film_year = suppressWarnings(min(as.integer(Ceremony_Year))), .groups = "drop")
```

#combine film year back to data
```{r}
combined2 <- combined %>%
  left_join(film_year, by = c("Film", "director"))
```

#filter acting categories 
```{r}
acting_aa <- AA_final %>%
  filter(str_detect(Category, "^ACTOR|^ACTRESS"))

actor_firsts <- acting_aa %>%
  group_by(Nominee) %>%
  summarise(
    first_nom_year = min(Ceremony_Year[Award_Status %in% c("Nominee", "Winner")], na.rm = TRUE),
    first_win_year = ifelse(any(Award_Status == "Winner"),
                            min(Ceremony_Year[Award_Status == "Winner"], na.rm = TRUE),
                            NA_integer_),
    .groups = "drop"
  )
```

#mark each credit as "prioir nominee/winner"
```{r}
credits <- combined2 %>%
  left_join(actor_firsts, by = c("Actor" = "Nominee")) %>%
  mutate(
    film_year = as.integer(film_year),
    prior_nominee = !is.na(first_nom_year) & !is.na(film_year) & first_nom_year < film_year,
    prior_winner  = !is.na(first_win_year) & !is.na(film_year) & first_win_year < film_year
  )
```

#Director % of credits that are prior nominees/winner rates
```{r}
dir_prior_rates <- credits %>%
  filter(!is.na(film_year)) %>%          
  group_by(director) %>%
  summarise(
    n_credits = n(),
    prior_nom_count = sum(prior_nominee, na.rm = TRUE),
    prior_win_count = sum(prior_winner, na.rm = TRUE),
    prior_nom_rate = prior_nom_count / n_credits,
    prior_win_rate = prior_win_count / n_credits,
    .groups = "drop"
  ) %>%
  arrange(desc(prior_nom_rate))

dir_prior_rates
```
# top 20 directors' rates
```{r}
topN <- 20
top_dirs <- credits %>%
  count(director, sort = TRUE) %>%
  slice_head(n = topN) %>%
  pull(director)

dir_prior_rates %>% filter(director %in% top_dirs)
```

#Director/Actor pair analysis, top 25 rows
```{r}
pair_stats <- credits %>%
  filter(!is.na(film_year)) %>%
  group_by(director, Actor) %>%
  summarise(
    n_films_together = n_distinct(Film),
    any_prior_nominee = any(prior_nominee, na.rm = TRUE),
    any_prior_winner = any(prior_winner, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  filter(any_prior_nominee | any_prior_winner) %>%
  arrange(desc(n_films_together))

head(pair_stats, 25)
```
#director-actor pairings
```{r}
dir_repeat <- pair_stats %>%
  group_by(director) %>%
  summarise(
    n_prior_awarded_actors = n(),
    max_repeat_with_prior_awarded = max(n_films_together),
    .groups = "drop"
  ) %>%
  arrange(desc(max_repeat_with_prior_awarded))

dir_repeat
```
#Chi-Square test
```{r}
top_credits <- credits %>%
  filter(director %in% top_dirs, !is.na(film_year)) %>%
  mutate(prior_nominee = factor(prior_nominee, levels = c(FALSE, TRUE)))

tab_prior <- table(top_credits$director, top_credits$prior_nominee)
chisq.test(tab_prior, simulate.p.value = TRUE, B = 10000)
```
#Chi-Square by directors. Result: the rate of working with actors who were already nominated/winners differs significantly by director.
```{r}
my_dirs <- c(
  "Adam McKay",
  "Alejandro González Iñárritu",
  "Alexander Payne",
  "Aaron Sorkin",
  "Christopher Nolan",
  "Clint Eastwood",
  "Darren Aronofsky",
  "David Fincher",
  "David O. Russell",
  "Greta Gerwig",
  "James Mangold",
  "Martin Scorsese",
  "Paul Thomas Anderson",
  "Pedro Almodóvar",
  "Peter Jackson",
  "Quentin Tarantino",
  "Ridley Scott",
  "Ron Howard",
  "Sam Mendes",
  "Stephen Daldry",
  "Stephen Frears",
  "Steven Spielberg",
  "Woody Allen"
  
)

dat_sel <- credits %>%
  filter(director %in% my_dirs, !is.na(film_year)) %>%
  mutate(prior_nominee = factor(prior_nominee, levels = c(FALSE, TRUE)))

tab_sel <- table(dat_sel$director, dat_sel$prior_nominee)
tab_sel

chisq.test(tab_sel, simulate.p.value = TRUE, B = 10000)
```
# Which directors differ from tab_self. Large positive residuals (in the TRUE column) = more prior nominees than expected
```{r}
prior_rate <- tab_sel[, "TRUE"] / rowSums(tab_sel)
sort(prior_rate, decreasing = TRUE)
```
#which dirctors deviate
```{r}
chi <- chisq.test(tab_sel, simulate.p.value = TRUE, B = 10000)
stdres_true <- chi$stdres[, "TRUE"]
sort(stdres_true, decreasing = TRUE)
```
#Chi-square for all directors
```{r}
all_credits <- credits %>%
  filter(!is.na(film_year)) %>%
  mutate(prior_nominee = factor(prior_nominee, levels = c(FALSE, TRUE)))

tab_all <- table(all_credits$director, all_credits$prior_nominee)
tab_all

chisq.test(tab_all, simulate.p.value = TRUE, B = 10000)
```
```{r}
chi_all <- chisq.test(tab_all, simulate.p.value = TRUE, B = 10000)

stdres_true <- chi_all$stdres[, "TRUE"]
head(sort(stdres_true, decreasing = TRUE), 50)   
head(sort(stdres_true, decreasing = FALSE), 50)  

prior_rate <- tab_all[, "TRUE"] / rowSums(tab_all)
head(sort(prior_rate, decreasing = TRUE), 50)
```

